---
title: "Summary_Statistics"
author: "Weijia Zhang (Vega)"
date: "2/26/2017"
output: html_document
---
#### Summary Statistics

```{r, message = FALSE, warning= FALSE}
library(etl)
library(DBI)
library(citibike)
library(utils)
library(data.table)
library(lubridate)
library(ggplot2)
# database
# example local directory on desktop 
bikes <- etl("citibike", dir = "~/Desktop/citibike_data", db = src_mysql_cnf("citibike", groups = "scidb"))
```


```{r}
# if the data is not loaded, uncomment the following code
#bikes <- bikes %>%
#  etl_load(years = 2013:2016, months = 1:12)
summary(bikes)
# analysis by starttime
trip_starttime<- bikes %>%
  tbl("trips_sub") %>%
  mutate(year = YEAR(Start_Time), month = MONTH(Start_Time), day = DAY(Start_Time), hour = HOUR(Start_Time), minute = MINUTE(Start_Time)) 


```

#### Breakdown of usertype
```{r}
bikes %>%
  tbl("trips") %>%
  group_by(User_Type)%>%
  summarise(count = n())
  
```

#### Breakdown of gender
```{r}
bikes %>%
  tbl("trips") %>%
  group_by(Gender)%>%
  summarise(count = n())
```


#### How the start stations are being used?
```{r}
# by start station
start_station_use <- bikes %>%
  tbl("trips_sub") %>%
  group_by(Start_Station_ID) %>%
  summarise(num_use = n()) %>%
  collect()
```

#### How many unique stations by end station?
```{r}

```

#### What's the use distribution of different stations in general?
with all the data including round trips
```{r}
colnames(start_station_use)[1] <- "ID"
colnames(end_station_use)[1] <- "ID"
total_station_use <- full_join(start_station_use, end_station_use, by = "ID")
colnames(total_station_use)[2] <- "start_use"
colnames(total_station_use)[3] <- "end_use"
# replace NAs with 0
total_station_use[is.na(total_station_use)] <- 0
ggplot(data = total_station_use) + geom_point(aes(x = start_use, y = end_use, group = ID))+ geom_abline(intercept = 0, slope = 1, color = "Red")

```



#### What is the distribution of bike volume across stations?
```{r}
trip_volumne <- bikes %>%
  tbl("trips_sub") %>%
  group_by(Start_Station_ID, End_Station_ID) %>%
  summarise(trip_volume = n()) %>%
  arrange(desc(trip_volume)) %>%
  collect(n = Inf)
nrow(trip_volumne)
# get the round trip stations
trip_round <- trip_volumne %>%
  filter(Start_Station_ID == End_Station_ID) %>%
  select(Start_Station_ID, trip_volume)
colnames(trip_round)[1] <- "Round_ID"
# two way trips only
trip_twoway <- left_join(total_station_use, trip_round, by= c("ID" = "Round_ID"))
# replace NAs with 0
trip_twoway[is.na(trip_twoway)] <- 0
trip_twoway <- trip_twoway %>%
  mutate(start_use = start_use - trip_volume) %>%
  mutate(end_use = end_use - trip_volume)
ggplot(data = trip_twoway) + geom_point(aes(x = start_use, y = end_use, group = ID))+ geom_abline(intercept = 0, slope = 1, color = "Red")



# plot the trip volume distribution
boxplot(trip_volumne$trip_volume)
plot(density(trip_volumne$trip_volume))
# take out the outlier
plot(density(trip_volumne[-1,]$trip_volume))
boxplot(trip_volumne[-1,]$trip_volume)
```


#### Trip distribution per year
```{r}
trip_vol_year <- bikes %>%
  tbl("trips") %>%
  mutate(year = YEAR(Start_Time)) %>%
  group_by(year) %>%
  summarise(total_trip = n()) %>%
  collect()
ggplot(data = trip_vol_year, aes(year,total_trip))+ geom_bar(stat = "identity", fill ="Light Blue")
```

#### Trip distribution per month
```{r}
trip_vol_month <- bikes %>%
  tbl("trips") %>%
  mutate(year = YEAR(Start_Time), month = MONTH(Start_Time)) %>%
  group_by(year, month) %>%
  summarise(total_trip = n()) %>%
  collect()
ggplot(data = trip_vol_month, aes(month,total_trip))+ geom_line(color ="Light Blue") + facet_wrap(~ year)

```

#### Trip distribution per day
```{r}
trip_vol_day <- bikes %>%
  tbl("trips") %>%
  mutate(year = YEAR(Start_Time), month = MONTH(Start_Time), day = DAY(Start_Time)) %>%
  group_by(year, month, day) %>%
  summarise(total_trip = n()) %>%
  collect()
ggplot(data = trip_vol_day, aes(day,total_trip))+ geom_line(color ="Light Blue") + facet_grid(month~ year)
```


